A glocal approach to designing RNA sequences with target secondary structure and nucleotide distribution

### 2013-01-08 ISMB/introduction_ISMB.tex
update abstract

In this paper we introduce \RNApyro, a RNA secondary structure design algorithm that benefits of our recent algorithmic advances \cite{Reinharz:2013aa} to expand our original \RNAensign algorithm \cite{Levin:2012kx}. \RNApyro addresses previous limitations of \RNAensign and offers new functionalities. First, while our previous program had a running time complexity of $\mathcal{O}(n^4)$, \RNApyro has now a linear-time and space complexity that allows him to demonstrate similar speeds to any local search algorithm. Next, \RNApyro is \textit{seed-less}. Unlike \RNAensign, it does not require a seed sequence to start its search. Finally, \RNApyro implements a novel algorithm using weighted sampling techniques \cite{Bodini2010} that enables us to control, for the first time, \textit{explicitly} the \texttt{C+G}-content of the solution. This functionality is essential because wild sequences within living organisms often present medium or low \texttt{C+G}-content. Presumably to offer better transcription rates and/or structural plasticity. Previous programs do not allow to control this parameter and tend to output sequences with high \texttt{C+G}-contents.

We demonstrate the performance of our algorithms on a set of real RNA structures with sizes ranging from *** to *** nucleotides. To complete this study, we develop an hybrid method combining our global sampling approach with local search strategies such as the one implemented in \RNAinverse. Remarkably, our glocal methodology overcome both local and global approaches.

>>>>

In this paper we introduce \ourprog, a RNA secondary structure design algorithm that benefits of our recent algorithmic advances \cite{Reinharz:2013aa} to expand our original \RNAensign algorithm \cite{Levin:2012kx}. \ourprog addresses previous limitations of \RNAensign and offers new functionalities. First, while our previous program had a running time complexity of $\mathcal{O}(n^4)$, \ourprog has now a linear-time and space complexity that allows him to demonstrate similar speeds to any local search algorithm. Next, \ourprog is \textit{seed-less}. Unlike \RNAensign, it does not require a seed sequence to start its search. Finally, \ourprog implements a novel algorithm using weighted sampling techniques \cite{Bodini2010} that enables us to control, for the first time, \textit{explicitly} the \gc~content of the solution. This functionality is essential because wild sequences within living organisms often present medium or low \gc~content. Presumably to offer better transcription rates and/or structural plasticity. Previous programs do not allow to control this parameter and tend to output sequences with high \gc~contents.

We demonstrate the performance of our algorithms on a set of real RNA structures extracted from the \textsf{RNA STRAND} database \cite{andronescu2008rna}. To complete this study, we develop an hybrid method combining our global sampling approach with local search strategies such as the one implemented in \RNAinverse. Remarkably, our glocal methodology overcome both local and global approaches.

### 2013-01-08 ISMB/main_ISMB.tex
algo
Method... interrupted
Further into methods section
res / supp data
Some typos + figs
update abstract

\documentclass[11pt]{article}
\usepackage[top=1in, bottom=1in,right=1in,left=1in]{geometry}                % See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   % ... or a4paper or a5paper or ... 
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm} %pour les newtheorem
\usepackage{epstopdf}
\usepackage{xspace}
\usepackage{relsize}
\usepackage[numbers]{natbib}
\usepackage[colorinlistoftodos]{todonotes}

@@

\newcommand{\ourprog}{\texttt{RNA-MoIP}\xspace}

@@

\newcommand{\RNAmutants}{\texttt{RNAmutants}\xspace}
\newcommand{\RNApyro}{\texttt{RNApyro}\xspace}

\newcommand{\red}[1]{{\color{red}#1}}
\newcommand{\farna}{\texttt{FARNA}\xspace}
\newcommand{\mcfoldmcsym}{\texttt{MC-Pipeline}\xspace}
\newcommand{\mcfold}{\texttt{MC-Fold}\xspace}
\newcommand{\mcsym}{\texttt{MC-Sym}\xspace}
\newcommand{\nast}{\texttt{NAST}\xspace}
\newcommand{\ifoldrna}{\texttt{iFoldRNA}\xspace}
\newcommand{\rnafold}{\texttt{RNAfold}\xspace}
\newcommand{\rnasubopt}{\texttt{RNAsubopt}\xspace}
\newcommand{\rnawolf}{\texttt{RNAwolf}\xspace}
\newcommand{\rnastructure}{\texttt{RNAstructure}\xspace}
\newcommand{\contrafold}{\texttt{contrafold}\xspace}
\newcommand{\unafold}{\texttt{unafold}\xspace}
\newcommand{\rnadd}{\texttt{RNA2D3D}\xspace}
\newcommand{\assemble}{\texttt{assemble}\xspace}
\newcommand{\fred}{\texttt{FR3D}\xspace}
\newcommand{\rnajunction}{\texttt{RNAjunction}\xspace}
\newcommand{\rnamotif}{\texttt{RNAmotif}\xspace}
\newcommand{\treefolder}{\texttt{TreeFolder}\xspace}
\newcommand{\barnacle}{\texttt{BARNACLE}\xspace}
\newcommand{\contextfold}{\texttt{contextfold}\xspace}
%\newcommand{\citep}{\cite}
\newcommand{\Z}[3]{\mathcal{Z}_{\substack{(#1)\\ [#3]}}^{#2}}
\newcommand{\Y}[3]{\mathcal{Y}_{\substack{(#1)\\ [#3]}}^{#2}}
\newcommand{\B}{\mathcal{B}}
\newcommand{\Kron}{\delta}
\newcommand{\ub}{\bullet}

\newcommand{\Ab}{{\sf{A}}}
\newcommand{\Cb}{{\sf{C}}}
\newcommand{\Gb}{{\sf{G}}}
\newcommand{\Ub}{{\sf{U}}}

\title{}
\author{}
\date{}

@@

\begin{document}
\maketitle
%\section{}
%\subsection{}
\begin{abstract}

\noindent
\textbf{Key words:} RNA, mutations, secondary structure
\end{abstract}

@@

>>>>

\documentclass[11pt]{article}
\usepackage[top=1in, bottom=1in,right=1in,left=1in]{geometry}                % See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   % ... or a4paper or a5paper or ... 
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm} %pour les newtheorem
\usepackage{epstopdf}
\usepackage{xspace}
\usepackage{tcolorbox}
\usepackage{relsize}
\usepackage[numbers]{natbib}
\usepackage[colorinlistoftodos]{todonotes}

@@

\usepackage[noend,ruled,vlined]{algorithm2e}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}

\newcommand{\ourprog}{\texttt{Incarnation}\xspace}

@@

\newcommand{\RNAmutants}{\texttt{RNAmutants}\xspace}
\newcommand{\RNApyro}{\texttt{RNApyro}\xspace}
\newcommand{\RNAinverse}{\texttt{RNAinverse}\xspace}
\newcommand{\RNASSD}{\texttt{RNA-SSD}\xspace}
\newcommand{\INFORNA}{\texttt{INFO-RNA}\xspace}
\newcommand{\NUPACK}{\texttt{NUPACK}\xspace}
\newcommand{\RNAiFOLD}{\texttt{RNAiFOLD}\xspace}
\newcommand{\frankenstein}{\texttt{Frnakenstein}\xspace}
\newcommand{\RNAexinv}{\texttt{RNAexinv}\xspace}
\newcommand{\rnaDesign}{\texttt{rnaDesign}\xspace}
\newcommand{\RNAensign}{\texttt{RNA-ensign}\xspace}
\newcommand{\GCContent}{\Gb\Cb-content\xspace}
\newcommand{\SFold}{\texttt{SFold}\xspace}

\newcommand{\red}[1]{{\color{red}#1}}
\newcommand{\farna}{\texttt{FARNA}\xspace}
\newcommand{\mcfoldmcsym}{\texttt{MC-Pipeline}\xspace}
\newcommand{\mcfold}{\texttt{MC-Fold}\xspace}
\newcommand{\mcsym}{\texttt{MC-Sym}\xspace}
\newcommand{\nast}{\texttt{NAST}\xspace}
\newcommand{\ifoldrna}{\texttt{iFoldRNA}\xspace}
\newcommand{\rnafold}{\texttt{RNAfold}\xspace}
\newcommand{\rnasubopt}{\texttt{RNAsubopt}\xspace}
\newcommand{\rnawolf}{\texttt{RNAwolf}\xspace}
\newcommand{\rnastructure}{\texttt{RNAstructure}\xspace}
\newcommand{\contrafold}{\texttt{contrafold}\xspace}
\newcommand{\unafold}{\texttt{unafold}\xspace}
\newcommand{\rnadd}{\texttt{RNA2D3D}\xspace}
\newcommand{\assemble}{\texttt{assemble}\xspace}
\newcommand{\fred}{\texttt{FR3D}\xspace}
\newcommand{\rnajunction}{\texttt{RNAjunction}\xspace}
\newcommand{\rnamotif}{\texttt{RNAmotif}\xspace}
\newcommand{\treefolder}{\texttt{TreeFolder}\xspace}
\newcommand{\barnacle}{\texttt{BARNACLE}\xspace}
\newcommand{\contextfold}{\texttt{contextfold}\xspace}
%\newcommand{\citep}{\cite}
\newcommand{\Z}[2]{\mathcal{Z}{\substack{(#1)\\ [#2]}}}
\newcommand{\Y}[3]{\mathcal{Y}_{\substack{(#1)\\ [#3]}}^{#2}}
\newcommand{\B}{\mathcal{B}}
\newcommand{\Kron}{\delta}
\newcommand{\ub}{\bullet}

\newcommand{\Ab}{{\sf{A}}\xspace}
\newcommand{\Cb}{{\sf{C}}\xspace}
\newcommand{\Gb}{{\sf{G}}\xspace}
\newcommand{\Ub}{{\sf{U}}\xspace}

\newcommand{\gc}{{\#\text{{\sf gc}}}}
\newcommand{\PE}[1]{E(#1)}
\newcommand{\EI}{\text{EI}}
\newcommand{\ES}{\text{{\sf Es}}}
\newcommand{\ISO}{\text{ISO}}
\newcommand{\Prob}{\text{P}}
\newcommand{\Target}{\text{\sf S}}
\newcommand{\CNBP}[1]{\gamma(#1)}

\newcommand{\CFramed}[2][gray]{\begin{tcolorbox}[colframe=#1,title=To be included only if space allows\ldots,colback=white]#1\end{tcolorbox}}
\renewcommand{\CFramed}[2][gray]{}

\title{A glocal approach to designing RNA sequences with target secondary structure and nucleotide distribution}
\author{Vladimir Reinharz$^1$, Yann Ponty$^{2,*}$, J\'er\^{o}me Waldisp\"{u}hl$^{1,*}$}
\date{$^1$ School of Computer Science, McGill University, Montreal, Canada, $^2$ Laboratoire d'informatique, ƒcole Polytechnique, Palaiseau, France.\\Corresponding authors: \texttt{jeromew@cs.mcgill.ca}, \texttt{yann.ponty@lix.polytechnique.fr}}

@@

\begin{document}
\maketitle
%\section{}
%\subsection{}
\begin{abstract}
\noindent
\textbf{Motivations:} The design of RNA sequences folding into predefined secondary structures is a milestone for many synthetic biology and gene therapy studies. Most of the current software use similar local search strategies (i.e. a random seed is progressively adapted to acquire the desired folding propertues) and more importantly do not allow the user to control explicitly the nucleotide distribution such as the \gc~content in their sequences. However, the latter is an important criteria for large-scale applications as it could presumably be used to design sequences with better transcription rates and/or structural plasticity. We recently introduced a novel global sampling approach for the RNA secondary structure design problem.

\noindent
\textbf{Results:}
In this paper, we introduce \ourprog

\noindent
\textbf{Key words:} RNA, secondary structure, design, weighted sampling, GC-content.
\end{abstract}

@@

\input{supplementary_ISMB.tex}
